---
title: "Foundations from Advanced R by Hadley Wickham"
author: "Amanda D. Clark"
output: html_notebook

---

# Intro
"The goal is to improve my R programming skills, I will delve deeper into the magic behind the R tools (packages, functions, etc.) that I used frequently. I hope to build a better sense of developing user-friendly programming using best practices (FAIR).

This notebook contains logic, examples, and quiz answers from the "Foundations" section of (Advanced R by Hadley Wickham)[https://adv-r.hadley.nz/index.html]. 


## Names and values

What is the distinction between objects and their names in R? Knowing this information will help make decisions about how to write faster code by avoiding unnecessary/accidental copies and memory usage. When will R copy an object? This means knowing the distinction between names and values.

### Pre-Quiz

```{r}
##1
df <- data.frame(runif(3), runif(3))
names(df) <- c(1, 2)

df <- df %>% mutate('3' = sum(df$'1', df$'2'))

##2 

x <- runif(1e6)
y <- list(x, x, x) #24MB


##3
a <- c(1, 5, 3, 2)
b <- a # copied here
b[[1]] <- 10
```

Okay, so that pre-quiz wasn't that easy. ##1 I'm sure that my third column is not correct in the context of being a sum across 1 & 2 for 3. Also the fact that the column names are numbers is a real bummer, because they have to be '' quoted to be interpreted correctly. 
The answers to the other questions are as commented in the code block. 


### Binding basics

Consider:

```{r}
x <- c(1,2,3)
```

Inaccurate pseudocode: "create an object named x, containing the values 1,2,3"
Accurate pseudocode: "creating an object, specifically a vector of values [c(1,2,3)] and binding that object to a name [x]

This means the name is a reference value. Say if we said:

```{r}
y <- x 
```

Here we aren't making another copy of the values, you are making another binding to those values. This means my answer in ##3 is incorrect when it asked which line it was copying on... 
What really happened was 'b' is a new binding to the values that 'a' was bound to first. 

Outside of bindings, objects have addresses, or where they are in memory, called an identifier. You can check an object's identifier with `lobstr::obj_addr()`
Let's use this function to determine if x and y are bound to the same identifier:

```{r}
obj_addr(x)
obj_addr(y)
```

Prior to running the previous code block, x & y had different identifiers. After running the code block, they became the same.
Once I restart and rerun, the object identifier will be different. 

#### Non-syntactic names

R has established rules for what is considered a valid name. A syntactic name has letters, digits, `.`, and `_`, but it cannot start with `_` or digits. You also should not be using reserved words like function or TRUE or NULL (use `?Reserved` to see full list). These would all be considered nonsyntactic names, and will produce errors:

```{r}
_adc <- 1

if <- 10
```

If you really want to use them, you can override the rules with backticks or quotes*...but you really shouldn't do this. Unfortunately, that won't stop other people who may do this, so you should know how to handle it:

```{r}
`_abc` <- 1
`_abc`

"_abc" <- 1


```
* the caveat with assigning with quotes is that you have to retrieve the bound value using a different syntax. 

#### Exercises

1. Explain the relationship between a, b, c and d in the following code:

```{r}
a <- 1:10
b <- a
c <- b
d <- 1:10
```

a and d are copies of one another. 
b and c are different bindings to the same object that a is bound to. 

2. The following code accesses the mean function in multiple ways. Do they all point to the same underlying function object? Verify this with `lobstr::obj_addr()`.

```{r}
mean
base::mean
get("mean")
evalq(mean)
match.fun("mean")
```

I'm gonna say yes, because they all share the same bytecode? After verification with `lobstr` It is confirmed (but the bytecode did not match the id obtained from `lobstr`)! 

3. By default, base R data import functions, like read.csv(), will automatically convert non-syntactic names to syntactic ones. Why might this be problematic? What option allows you to suppress this behaviour?

This may be problematic because we might get overlapping names? could you silence an option? After using `?read.csv` you can set `check.names` to `FALSE`! It is changed by make.names and usually ensures that there are no duplicates so ignore what I said earlier. Maybe the issue would come in with reproducibility and having to update the code everytime because the renames might be different. 

4. What rules does make.names() use to convert non-syntactic names into syntactic ones?

"The character "X" is prepended if necessary. All invalid characters are translated to ".". A missing value is translated to "NA". Names which match `R` keywords have a dot appended to them. Duplicated values are altered by `make.unique`."

5. I slightly simplified the rules that govern syntactic names. Why is .123e1 not a syntactic name? Read ?make.names for the full details.

It's not syntactic because there is a number following the `.` and not a letter. 

### Copy-on-modify

Consider the following code:

```{r}
x <- c(1, 2, 3)
y <- x

y[[3]] <- 4
x
#> [1] 1 2 3
```

Here we are binding x and y to the same value, but then we modify y. This modification doesn't effect x though, which means these are no longer bound to the same object 


