---
title: "Foundations from Advanced R by Hadley Wickham"
output: html_notebook
--- 


# Intro 

The goal is to improve my `R` programming skills, I will delve deeper into the magic behind the R tools (packages, functions, etc.) that I used frequently. I hope to build a better sense of developing user-friendly programming using best practices (FAIR).

This notebook contains logic, examples, and quiz answers from the "Foundations" section of [Advanced R by Hadley Wickham](https://adv-r.hadley.nz/index.html). 


## Names and values

What is the distinction between objects and their names in R? Knowing this information will help make decisions about how to write faster code by avoiding unnecessary/accidental copies and memory usage. When will R copy an object? This means knowing the distinction between names and values.

### Pre-Quiz

```{r}
##1
df <- data.frame(runif(3), runif(3))
names(df) <- c(1, 2)

df <- df %>% mutate('3' = sum(df$'1', df$'2'))

##2 

x <- runif(1e6)
y <- list(x, x, x) #24MB


##3
a <- c(1, 5, 3, 2)
b <- a # copied here
b[[1]] <- 10
```

Okay, so that pre-quiz wasn't that easy. ##1 I'm sure that my third column is not correct in the context of being a sum across 1 & 2 for 3. Also the fact that the column names are numbers is a real bummer, because they have to be '' quoted to be interpreted correctly. 
The answers to the other questions are as commented in the code block. 


### Binding basics

Consider:

```{r}
x <- c(1,2,3)
```

Inaccurate pseudocode: "create an object named x, containing the values 1,2,3"
Accurate pseudocode: "creating an object, specifically a vector of values [c(1,2,3)] and binding that object to a name [x]

This means the name is a reference value. Say if we said:

```{r}
y <- x 
```

Here we aren't making another copy of the values, you are making another binding to those values. This means my answer in ##3 is incorrect when it asked which line it was copying on... 
What really happened was 'b' is a new binding to the values that 'a' was bound to first. 

Outside of bindings, objects have addresses, or where they are in memory, called an identifier. You can check an object's identifier with `lobstr::obj_addr()`
Let's use this function to determine if x and y are bound to the same identifier:

```{r}
obj_addr(x)
obj_addr(y)
```

Prior to running the previous code block, x & y had different identifiers. After running the code block, they became the same.
Once I restart and rerun, the object identifier will be different. 

#### Non-syntactic names

R has established rules for what is considered a valid name. A syntactic name has letters, digits, `.`, and `_`, but it cannot start with `_` or digits. You also should not be using reserved words like function or TRUE or NULL (use `?Reserved` to see full list). These would all be considered nonsyntactic names, and will produce errors:

```{r}
_adc <- 1

if <- 10
```

If you really want to use them, you can override the rules with backticks or quotes*...but you really shouldn't do this. Unfortunately, that won't stop other people who may do this, so you should know how to handle it:

```{r}
`_abc` <- 1
`_abc`

"_abc" <- 1


```
* the caveat with assigning with quotes is that you have to retrieve the bound value using a different syntax. 

#### Exercises

1. Explain the relationship between a, b, c and d in the following code:

```{r}
a <- 1:10
b <- a
c <- b
d <- 1:10
```

a and d are copies of one another. 
b and c are different bindings to the same object that a is bound to. 

2. The following code accesses the mean function in multiple ways. Do they all point to the same underlying function object? Verify this with `lobstr::obj_addr()`.

```{r}
mean
base::mean
get("mean")
evalq(mean)
match.fun("mean")
```

I'm gonna say yes, because they all share the same bytecode? After verification with `lobstr` It is confirmed (but the bytecode did not match the id obtained from `lobstr`)! 

3. By default, base R data import functions, like read.csv(), will automatically convert non-syntactic names to syntactic ones. Why might this be problematic? What option allows you to suppress this behaviour?

This may be problematic because we might get overlapping names? could you silence an option? After using `?read.csv` you can set `check.names` to `FALSE`! It is changed by make.names and usually ensures that there are no duplicates so ignore what I said earlier. Maybe the issue would come in with reproducibility and having to update the code everytime because the renames might be different. 

4. What rules does make.names() use to convert non-syntactic names into syntactic ones?

"The character "X" is prepended if necessary. All invalid characters are translated to ".". A missing value is translated to "NA". Names which match `R` keywords have a dot appended to them. Duplicated values are altered by `make.unique`."

5. I slightly simplified the rules that govern syntactic names. Why is .123e1 not a syntactic name? Read ?make.names for the full details.

It's not syntactic because there is a number following the `.` and not a letter. 

### Copy-on-modify

Consider the following code:

```{r}
x <- c(1, 2, 3)
y <- x

y[[3]] <- 4
x
#> [1] 1 2 3
```

Here we are binding x and y to the same value, but then we modify y. This modification doesn't effect x though, which means these are no longer bound to the same object and we have created a copy of the original values... so in the pre-quiz, the answer is actually the last line of the code. 

it seems like r objects are technically immutable, but writer refrains due to prominent exceptions. 

#### `tracemem()`

You can track when an object gets copied using `tracemem()`: 

```{r}
x <- c(1, 2, 3)
cat(tracemem(x), "\n")
```

So now I should be able to run the following two lines and see when `x` gets copied. It shouldn't happen from running the second line, but `tracemem()` should show me that a copy was made when I run the second line: 

```{r}
y <- x #`tracemem()` did nothing when I ran this line alone
y[[3]] <- 4L #`tracemem()` responded to this line 

```
Sweet. I expect that if I modify y again, `tracemem()` will not respond because I am editing a different object with only 1 binding and not making a copy in any way. I also expect that it will have the object identifier that is on the right of the arrow from the previous output. 

```{r}
y[[3]] <- 5L

untracemem(x)
```

Okay so `tracemem()` still responded, and I'm not sure why... {revisit}
Turn off tracing with the second line in the chunk above!

#### Function calls

This copy-on-modify extends in functions as well. Consider this code:

```{r}
f <- function(a) {
  a
}

x <- c(1, 2, 3)
cat(tracemem(x), "\n")


z <- f(x)
cat(tracemem(z), "\n")

untracemem(x)
untracemem(z)
```

We have a function that takes something and does not modify it at all. If we put x in this function, we can expect that x and z (the name bound to the output of the value f(x)) are bound to the same thing (this binding is because a from the function is bound as well). If we were to modify x in the function, a copy would need to be created and z would be bound to a different copy of the original object. 

I added a second `cat(tracemem(), "\n")` argument for z to compare what happens to the id when the function makes modifications and when it doesn't. as you can imagine, the IDs were identical when the function did not modify the input. 

#### Lists

We can expect this behavior from lists as well (because elements of a list point to values), but there are some nuances:

```{r}
l1 <- list(1,2,3)
```

Here instead of storing the values, lists store references to those values (index?). l1 binds to the list object, but elements in the list object are stored references to the actual values? Let's make a copy and a modification to see ow things change with `lobstr::ref()`

```{r}
l2 <- l1
ref(l1, l2)
```
We see here that everything between the lists are shared. This is again because l1 & l2 are bound to the same list at the moment. What happens after modification?

```{r}
l2[[3]] <- 4
ref(l1, l2)
```

This time we get two trees and the list itself as well as the third index have different memory addresses now. Note that index 1 & 2 are shown a second time but with less information and are a different color. Lists undergo "shallow" copying, meaning the object and bindings are copied over, but the values being pointed to are not. These images were helpful in understanding this:
[Pre copy-on-modify](https://d33wubrfki0l68.cloudfront.net/52bc0e3da3382cba957a9d83397b6c9200906ce2/c72aa/diagrams/name-value/l-modify-1.png)
[Post copy-on-modify](https://d33wubrfki0l68.cloudfront.net/b844bb5a3443e1344299627f5760e2ae3a9885b5/e1c76/diagrams/name-value/l-modify-2.png)